Short-time simulations of open collective quantum spin systems

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
renamed the method to 'truncated short-time (TST) expansion' added conclusion and acknowledgments

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative operator $T^k$ are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\O\p{t}}$ to a fixed order $M$.  Note that using the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
qof matrix elements $T_{\v m\v n}$ that we may need to explicitly
compute roughly in half.

In principle, the factorial suppression of terms at higher orders of
the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
given time $t$, there exists a truncation order $M_t$ for which the
truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
accessible with limited computational resources, such that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
allows computation of the expectation value $\bk{\O\p{t}}$ to a
maximal time $t_{M_{\t{max}}}$.  As we will see in the following
section, $M_{\t{max}}=35$ will suffice for the computation of
collective spin correlators up to times that are relevant e.g.~for
spin squeezing protocols.

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\x,S_\y,S_\z}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of the coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
these exact and quantum trajectory simulations will be used to
benchmark the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics
  \label{fig:squeezing_coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$
  \label{fig:squeezing_incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:squeezing_coherent} coherent and
    \sref{fig:squeezing_incoherent} incoherent evolution, computed
    using exact methods (solid lines), quantum trajectory simulations
    (dots), and the truncated Taylor expansion in
    Eq.~\eqref{eq:order_truncation} with $M=35$ (darker dashed lines).
    Solid circles mark the times at which the truncated Taylor
    expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
as dynamics in the presence of spontaneous decay, excitation, and
dephasing of individual spins at rates $\chi$ (Figure
\ref{fig:squeezing_incoherent}), respectively described by the sets of
jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
appropriate transformations of the Hamiltonian and jump operators);
the only effect of this rotation on the results presented in Figure
\ref{fig:benchmarking} is to prolong the time for which the truncated
Taylor expansion agrees with the benchmarking simulations.  The reason
for different results in rotated basis has to do with the breakdown of
the truncated Taylor expansion, which we discuss further below.

The main lesson from Figure \ref{fig:benchmarking} is that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
essentially exact results right up until a sudden and drastic
departure that is simple to diagnose by inspection.  This drastic
departure occurs because the truncated Taylor expansion neglects
high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
contributions to a Heisenberg operator $\O\p{t}$ of interest
eventually become non-negligible.  For sufficiently large times $t$,
these individual contributions may generally be large compared to the
value of $\bk{\O\p{t}}$, in which case their truncation will yield
non-physical results.  This breakdown mechanism is also the reason for
prolonged agreement between the truncated Taylor expansion and
benchmarking methods when squeezing the initial state $\ket{-\Z}$
rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
initial-time correlators $\bk{\S_{\v m}}_{t=0}$ vanish unless
$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
substantially smaller number of neglected terms with non-zero
contribution to the correlator $\bk{\O\p{t}}$.

Although the truncated Taylor expansion breaks down at short times, it
has two key advantages over other methods to compute collective spin
correlators.  First, computing spin correlators with the truncated
Taylor expansion is generally much faster and requires much fewer
computing resources than the alternatives.  The quantum trajectory
Monte Carlo simulations performed for Figure
\ref{fig:squeezing_incoherent}, for example, take $\sim10^4$ CPU hours
to compute on standard modern computing hardware; the bulk of this
time is spent performing sparse matrix multiplication, leaving little
room to further optimize run time.  Parallelization can reduce the
actual run time of these quantum trajectory simulations to $\sim10$
hours, but at the cost of greatly increasing computing resource
requirements.  The truncated Taylor expansion results in Figure
\ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$ seconds to
compute with a single CPU on similar computing hardware.

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
spontaneous decay, excitation, and dephasing of individual spins at
rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that
TNT is much more robust to this kind of decoherence than OAT or TAT, a
finding that could not be deduced from the weak-decoherence
simulations presented in Figure \ref{fig:benchmarking}.
Strong-decoherence computations of this sort were used to put lower
bounds on theoretically achievable spin squeezing via TAT with
decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the truncated Taylor expansion and the
collective-spin structure constants calculated in this work.

\begin{figure}
  \centering
  \includegraphics{decoherence_strong.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under incoherent evolution spontaneous with decay, excitation, and
    dephasing of individual spins at rates
    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the
    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
    $M=35$.  Solid circles mark the times at which the truncated
    Taylor expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:decoherence_strong}
\end{figure}

@@

As a final example of collective-spin physics that is numerically
accessible via the truncated Taylor expansion of Heisenberg operators,
we consider the calculation of collective-spin two-time correlation
functions.  In particular, we consider the effect of decoherence on
the short-time behavior of the two-time connected correlator
\begin{align}
  C\p{t}
  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
  \label{eq:two_time}
\end{align}
in the context of the squeezing models in Section \ref{sec:squeezing}.
In an equilibrium setting, correlation functions of this sort contain
information about the linear response of Heisenberg operators to
perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
setting, they contribute to short-time linear response (see Appendix
\ref{sec:linear_response}).  Similar correlators have made appearances
as order parameters that diagnose time-crystalline phases of
matter\cite{tucker2018shattered}.  Related multi-time correlators,
known as out-of-time-ordered correlators, are commonly examined for
signatures of quantum chaos and information
scrambling[CITE]\cite{CITE}.

@@

Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
spins, initially in the state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
with and without various sources of decoherence.  The two-time
correlator $C\p{t}$ degrades at short times under certain types of
decoherence, but seems relatively stable to collective dephasing (with
jump operator $S_\z$), and can qualitatively change behavior under in
the presence of single-spin or collective excitation (with jump
operators $s_+^{(j)}$ or $S_+$).  While we do not have an operational
understanding of the correlations captured by $C\p{t}$ in the context
of spin squeezing, Figure \ref{fig:two_time} serves as an example for
the type of quantity and behavior that is accessible at short times
with the truncated Taylor expansion.  This example is straightforward
to extend to multi-time and out-of-time-ordered correlators, as well
as equilibrium settings.

@@

% extension to spin-boson and cavity systems / dicke model / chaos?
% different initial conditions <--> steady states or pseudo-steady states in models with a large separation of scales
% OTOCs; theoretical interest, scrambling; benchmarking "practical" OTOCs (with time-reversal of only coherent dynamics) against "true" OTOCs

>>>>

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative operator $T^k$ are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
We refer Eq.~\eqref{eq:order_truncation} as the truncated short-time
(TST) expansion of Heisenberg operators.  Unlike the truncation in
Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in
Eq.~\eqref{eq:order_truncation} tell us which operators are relevant
for computing the expectation value $\bk{\O\p{t}}$ to a fixed order
$M$.  Note that using the relation $\S_{\v m}^\dag=\S_{\v m^*}$ for
$\v m^*\equiv\p{m_-,m_\z,m_+}$, which by Hermitian conjugation of
Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
qof matrix elements $T_{\v m\v n}$ that we may need to explicitly
compute roughly in half.

In principle, the factorial suppression of terms at higher orders of
the expansion in Eq.~\eqref{eq:time_series} implies for any given time
$t$, there exists a truncation order $M_t$ for which the truncation
error in Eq.~\eqref{eq:order_truncation} is negligibly small.  In
practice, only a maximal truncation order $M_{\t{max}}$ is accessible
with limited computational resources, such that the TST expansion in
Eq.~\eqref{eq:order_truncation} only allows computation of the
expectation value $\bk{\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.
As we will see in the following section, $M_{\t{max}}=35$ will suffice
for the computation of collective spin correlators up to times that
are relevant e.g.~for spin squeezing protocols.

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\x,S_\y,S_\z}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of the coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
these exact and quantum trajectory simulations will be used to
benchmark the TST expansion in Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics
  \label{fig:squeezing_coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$
  \label{fig:squeezing_incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:squeezing_coherent} coherent and
    \sref{fig:squeezing_incoherent} incoherent evolution, computed
    using exact methods (solid lines), quantum trajectory simulations
    (dots), and the TST expansion in Eq.~\eqref{eq:order_truncation}
    with $M=35$ (darker dashed lines).  Solid circles mark the times
    at which the TST expansion gives an unphysical result with
    $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the TST expansion in
Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
as dynamics in the presence of spontaneous decay, excitation, and
dephasing of individual spins at rates $\chi$ (Figure
\ref{fig:squeezing_incoherent}), respectively described by the sets of
jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
appropriate transformations of the Hamiltonian and jump operators);
the only effect of this rotation on the results presented in Figure
\ref{fig:benchmarking} is to prolong the time for which the TST
expansion agrees with the benchmarking simulations.  The reason for
different results in rotated basis has to do with the breakdown of the
TST expansion, which we discuss further below.

The main lesson from Figure \ref{fig:benchmarking} is that the TST
expansion gives essentially exact results right up until a sudden and
drastic departure that is simple to diagnose by inspection.  This
drastic departure occurs because the TST expansion neglects
high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
contributions to a Heisenberg operator $\O\p{t}$ of interest
eventually become non-negligible.  For sufficiently large times $t$,
these individual contributions may generally be large compared to the
value of $\bk{\O\p{t}}$, in which case their truncation will yield
non-physical results.  This breakdown mechanism is also the reason for
prolonged agreement between the TST expansion and benchmarking methods
when squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$:
for initial states $\ket{-\Z}$, all initial-time correlators
$\bk{\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
\ref{sec:initial_conditions}), so there is a substantially smaller
number of neglected terms with non-zero contribution to the correlator
$\bk{\O\p{t}}$.

Although the TST expansion breaks down at short times, it has two key
advantages over other methods to compute collective spin correlators.
First, computing spin correlators with the TST expansion is generally
much faster and requires much fewer computing resources than the
alternatives.  The quantum trajectory Monte Carlo simulations
performed for Figure \ref{fig:squeezing_incoherent}, for example, take
$\sim10^4$ CPU hours to compute on standard modern computing hardware;
the bulk of this time is spent performing sparse matrix
multiplication, leaving little room to further optimize run time.
Parallelization can reduce the actual run time of these quantum
trajectory simulations to $\sim10$ hours, but at the cost of greatly
increasing computing resource requirements.  The TST expansion results
in Figure \ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$
seconds to compute with a single CPU on similar computing hardware.

Second, the TST expansion enables computing spin correlators in
strong-decoherence regimes entirely inaccessible to other methods.  As
an example, Figure \ref{fig:decoherence_strong} shows squeezing of
$N=10^4$ spins initially in $\ket{\X}$, undergoing spontaneous decay,
excitation, and dephasing of individual spins at rates
$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that TNT is
much more robust to this kind of decoherence than OAT or TAT, a
finding that could not be deduced from the weak-decoherence
simulations presented in Figure \ref{fig:benchmarking}.
Strong-decoherence computations of this sort were used to put lower
bounds on theoretically achievable spin squeezing via TAT with
decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the TST expansion and the collective-spin
structure constants calculated in this work.

\begin{figure}
  \centering
  \includegraphics{decoherence_strong.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under incoherent evolution spontaneous with decay, excitation, and
    dephasing of individual spins at rates
    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the TST
    expansion in Eq.~\eqref{eq:order_truncation} with $M=35$.  Solid
    circles mark the times at which the TST expansion gives an
    unphysical result with $\xi^2<0$.}
  \label{fig:decoherence_strong}
\end{figure}

@@

As a final example of collective-spin physics that is numerically
accessible via the TST expansion of Heisenberg operators, we consider
the calculation of collective-spin two-time correlation functions.  In
particular, we consider the effect of decoherence on the short-time
behavior of the two-time connected correlator
\begin{align}
  C\p{t}
  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
  \label{eq:two_time}
\end{align}
in the context of the squeezing models in Section \ref{sec:squeezing}.
In an equilibrium setting, correlation functions of this sort contain
information about the linear response of Heisenberg operators to
perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
setting, they contribute to short-time linear response (see Appendix
\ref{sec:linear_response}).  Similar correlators have made appearances
as order parameters that diagnose time-crystalline phases of
matter\cite{tucker2018shattered}.  Related multi-time correlators,
known as out-of-time-ordered correlators, are commonly examined for
signatures of quantum chaos and information
scrambling[CITE]\cite{CITE}.

@@

Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
spins, initially in the state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
with and without various sources of decoherence.  The two-time
correlator $C\p{t}$ degrades at short times under certain types of
decoherence, but seems relatively stable to collective dephasing (with
jump operator $S_\z$), and can qualitatively change behavior or grow
in the presence of single-spin or collective excitation (with jump
operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time} serves as
an example for the type of quantity and behavior that is accessible at
short times with the TST expansion.  This example is straightforward
to extend to multi-time and out-of-time-ordered correlators (OTOCs),
as well as equilibrium settings.

@@

We have presented a new method for computing correlators in collective
spin systems at short times.  This method is based on truncating a
short-time expansion of Heisenberg operators, and can access
correlators on time scales that are relevant to metrological
applications such as spin squeezing.  In order to evaluate the
truncated short-time (TST) expansion of Heisenberg operators, we have
computed the structure constants of a collective spin operator
algebra, which we hope will empower future analytical and numerical
studies of collective spin systems.  Even though we considered only
non-equilibrium spin-squeezing processes in this work, our method can
be applied directly in an equilibrium setting, and is straightforward
to generalize to systems such as trapped ions and optical cavities
with collective spin-boson interactions.  In such contexts, our method
may be used to benchmark the short-time effects of decoherence, or
compute quantities such as out-of-time-ordered correlators (OTOCs)
that diagnose the onset of quantum chaos and information scrambling.

\section{Acknowledgments}

We acknowledge helpful discussions with Robert Lewis-Swan and Kris
Tucker.  This work is supported by the Air Force Office of Scientific
Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary
University Research Initiative (MURI) grant; the Defense Advanced
Research Projects Agency (DARPA) and Army Research Office (ARO) grant
W911NF-16-1-0576; the National Science Foundation (NSF) grant
PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute of
Standards and Technology (NIST).
